---
title: "Homework 3"
format:
  html: default
  pdf: default
editor: visual
---

# Base R and R Basics

HINT: Remember that you can get help on any function by typing `?`(function name). For instance, `?rnorm` gives help on the `rnorm()` function.

## Creating and naming variables

1.  Create a variable called `x` and use it to store the result of the calculation `(3*(4+2))`.

    **Answer (code):**

    ```{r}
    x = (3*(4+2))
    ```

2.  Calculate the product of `x` (from the above question) times π.

    **Answer (code):**

    ```{r}
    # YOUR CODE HERE
    ```

3.  Use the `getwd()` function to show your current working directory. Did you set the working directory as part of an R project, in the console, or some other way? Is it the same as if you write getwd() in the console? why might these be different?

    **Answer (code + words):**

    ```{r}
    getwd()
    ```

    **Answer (words):** *Write your answers here.*

## Vectors

1.  Use the `c()` function to create a vector of numbers.

    **Answer (code):**

    ```{r}
    # YOUR CODE HERE
    ```

2.  Use the `c()` function to create a vector of characters.

    **Answer (code):**

    ```{r}
    # YOUR CODE HERE
    ```

3.  Use the `:` implicit function to create a vector of integers from 1 to 10.

    **Answer (code):**

    ```{r}
    # YOUR CODE HERE
    ```

4.  Explain *why* the following code returns what it does. Also address whether you think this was a good decision on the part of the designers of R?

    ```{r}
    v1 <- 1:3
    v2 <- c(1:4)
    v1 + v2
    ```

    **Answer (words):** *Write answers here.*

5.  Explain what the following code does. It may be helpful to reference the answer to the previous question:

    ```{r}
    c(1, 5, 9) + 3
    ```

    **Answer (words):** *Write answers here.*

6.  Remove (delete) every variable in your workspace.

    **Answer (code):**

    ```{r}
    # YOUR CODE HERE
    ```

## Graphics

1.  Load the tidyverse package. **NOTE:** Be sure to use the chunk option `message=FALSE` to suppress the messages that tidyverse prints when loaded.

    **Answer (code):**

    ```{r, message=FALSE}
    library(tidyverse)
    ```

<!-- -->

2.  Recreate the visualization of `body_mass_g` to `flipper_length_mm`, from the penguins data set, that is shown in question 8 of section 1.2.5 of [R4DS](https://r4ds.hadley.nz/data-visualize).

    **Answer (code):**

    ```{r}
    # If needed:
    # install.packages("palmerpenguins")
    library(palmerpenguins)

    # YOUR CODE HERE
    ```

3.  Explain why each aesthetic is mapped at the level that it is (i.e., at the global level, in the `ggplot()` function call, or at the geom level, in the `geom_XXX()` function call). Note: A lot of different options will work, but some options are clearly better than others.

    **Answer (words):** *Write answers here.*

4.  **(New)** Make a **bar chart** showing the number of penguins in the dataset for each `species`.

    -   Hint: start with `ggplot(penguins, aes(x = species)) + ...`

    **Answer (code):**

    ```{r}
    # YOUR CODE HERE
    ```

    **Answer (words):** *In 1–3 sentences: what does this plot tell you about the dataset?*

5.  **(New)** Make a **density plot** of `body_mass_g`, with a separate curve for each `species`.

    -   Hint: `geom_density()` and map `color = species` (and optionally `fill = species` with some `alpha`).

    **Answer (code):**

    ```{r}
    # YOUR CODE HERE
    ```

    **Answer (words):** *Write here.*

# What to submit

Submit your completed `.qmd` file **and** the rendered output (HTML or PDF).
